Main Figure 2. Bacterial Communities and Function

### 2021-10-29 reports/rmarkdown/manuscript/Figure2.Rmd
Fig.2 Version 2

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

#################
# COLOR PALETTS #
#################
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
# fig.asp=.6
# Define Layout and set labels
# layout(matrix(c(2,3,4,5,      6,7,  11,12,13,14,
#                 1,1,1,1,      1,8,  11,12,13,14,
#                 1,1,1,1,      1,9,  11,12,13,14,
#                 1,1,1,1,      1,10, 11,12,13,14,
#                 15,15,15,15,  0,0,  11,12,13,14
#                 ) , 
#               ncol= 10,byrow = T),widths = c(1,1,1,1,1,1,1,2,.5,6),heights = c(1.5,1.2,1.2,1.2,1.2))
# layout.show(3)

layout(matrix(c(1,2,3,4,5,
                6,2,3,4,5,
                7,7,7,10,10,
                8,8,8,10,10,
                9,9,9,10,10),
              nrow = 5,ncol = 5,byrow = T), widths = c(1.9,.6,1,.2,2),
       heights = c(1.5,1.5,.7,.7,.7))
figlabels <- letters

#########################
# BACTERIAL COMMUNITIES #
#########################
bac_communities <- factor(paste0("BC",sprintf("%02d",bac_communities)))
gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
 set.seed(1)
lB <- layout_nicely(gB,niter=3000,start.temp=30)

par(mar=c(0,0,3.5,5))
plot( gB , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(bac_communities)] ,
      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
      vertex.size=10, #main="Bacterial Communities",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
title(main = "           Bacterial Communities- Louvain", line = 0.5, cex.main = 1)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = bact_pal,pt.cex = 1,xpd=T)

@@

# par(mar=c(1,1,1,1))
# for(j in sort(unique(bac_communities)) ){
#   empty_plot()
#   text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
#   title(main=j,cex.main=.5)
# }

#####################
# BACTERIAL PICRUST #
#####################
picrust_data <- round(picrust_data[ rowSums(picrust_data>5)>=10, ])
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

@@

#saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
#write.csv2( top , "../../../results/Picrust_EdgeR_top_dge.csv" )
write.xlsx(top, file=paste0("../../../results/DE/","dge_picrust_",gr,"_conf.xlsx"))

@@

source("../enrichment_function.R")
gmt_list <- fgsea::gmtPathways("../../../data/KEGG_pathways_to_KO.tsv")
kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
names(gmt_list) <- kegg_names [ match ( names(gmt_list), as.character(kegg_names[,1]) ) , 2 ]
gmt_list <- lapply(gmt_list , function(x){
  na.omit(ko_names[x])
})
res_list <- lapply( unique( gene_module[h$order] ),function(x){
  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  return(temp)
} )
res <- map(res_list, ~rownames_to_column(.x, var = "Terms")) %>% 
    set_names(., paste0("module_",  seq_along(res_list)))
write.xlsx(res, file=paste0("../../../results/ENRICHMENT/Heatmap/","KEGG","_enrichment_picrust_",gr,".xlsx"))

@@

par(mar=c(4,.1,.5,7))
barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,names.arg = "")
title(xlab="-log10(pvalue)", line=2)
abline(v=2,xpd=F,col="red",lty=2)
points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,xpd=T,
      labels = rev(gsub("\\(\\)","",names(pvalues)) ),xpd=T)
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,
      labels = rev(names(pvalues)),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )

################
# PICRUST UMAP #
################

# based on 108 samples used in manuscript:
par(mar=c(2.5,2.5,3,4)) #b,l,t,r
set.seed(1)
UMAP_picrust <- uwot::umap(t(named_picrust_data[,sample_use]),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[factor(metadata$Luminal_gr)],
     #main="Top Functional Profile- UMAP",
     pch=21,frame=F,axes=F)
title(main = "Top Functional Profile- UMAP", line = 0.5, cex.main = 1)
#mtext("UMAP1", side=1, line=0.7, cex=0.8)
#mtext("UMAP2", side=2, line=0.7, cex=0.8)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(metadata$Luminal_gr)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T, cex=1)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,1,2,1)) #b,l,t,r
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

barlist( data = rbind(WS=mL ,
                      counts_16S=xL),
         main = "Lactobacillus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(WS=mM ,
                      counts_16S=xM),
         main = "Mobiluncus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(WS=mG ,
                      counts_16S=xG),
         main = "Gardnerella",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

@@

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

layout(matrix(c(1,
                2,
                3),
              nrow = 3,ncol = 1,byrow = T), widths = c(1,1,1))

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(3,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
xM <- rowsum(x , grepl("Monbilicus",rownames(x)) )[2,] / sum(grepl("Monbilicus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Mobiluncus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardnerella",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

```

>>>>

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}

layout(matrix(c(1,1,2,2,2,
                1,1,3,3,3,
                1,1,4,4,4,
                5,6,7,8,9,
                5,6,7,8,10),
              nrow = 5,ncol = 5,byrow = T), widths = c(.6,1.5,.2,1.9,2),
       heights = c(.6,.6,.6,1,1.5))
figlabels <- letters

## A
#########################
# BACTERIAL COMMUNITIES #
#########################
bac_communities <- factor(paste0("BC",sprintf("%02d",bac_communities)))
gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
 set.seed(1)
lB <- layout_nicely(gB,niter=3000,start.temp=30)

par(mar=c(0,0,3.5,5))
plot( gB , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(bac_communities)] ,
      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
title(main = "           Bacterial Communities- Louvain", line = 0.5, cex.main = 1)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = bact_pal,pt.cex = 1,xpd=T)

@@

## B
###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,2,1)) #b,l,t,r
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

barlist( data = rbind(WS=mL ,
                      "16S counts"=xL),
         main = "Lactobacillus",
         genes = c("WS","16S counts"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(WS=mM ,
                      "16S counts"=xM),
         main = "Mobiluncus",
         genes = c("WS","16S counts"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(WS=mG ,
                      "16S counts"=xG),
         main = "Gardnerella",
         genes = c("WS","16S counts"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

## C
#####################
# BACTERIAL PICRUST #
#####################
picrust_data <- round(picrust_data[ rowSums(picrust_data>5)>=10, ])
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

@@

@@

source("../enrichment_function.R")
gmt_list <- fgsea::gmtPathways("../../../data/KEGG_pathways_to_KO.tsv")
kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
names(gmt_list) <- kegg_names [ match ( names(gmt_list), as.character(kegg_names[,1]) ) , 2 ]
gmt_list <- lapply(gmt_list , function(x){
  na.omit(ko_names[x])
})
res_list <- lapply( unique( gene_module[h$order] ),function(x){
  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  return(temp)
} )
res <- map(res_list, ~rownames_to_column(.x, var = "Terms")) %>% 
    set_names(., paste0("module_",  seq_along(res_list)))

@@

par(mar=c(4,.1,.5,7))
barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,names.arg = "")
title(xlab="-log10(pvalue)", line=2)
abline(v=2,xpd=F,col="red",lty=2)
points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,xpd=T,
      labels = rev(gsub("\\(\\)","",names(pvalues)) ),xpd=T)
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,
      labels = rev(names(pvalues)),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)

## D
################
# PICRUST UMAP #
################

# based on 108 samples used in manuscript:
par(mar=c(2.5,2.5,3,4)) #b,l,t,r
set.seed(1)
UMAP_picrust <- uwot::umap(t(named_picrust_data[,sample_use]),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[factor(metadata$Luminal_gr)],
     pch=21,frame=F,axes=F)
title(main = "Top Functional Profile- UMAP", line = 0.5, cex.main = 1)
#mtext("UMAP1", side=1, line=0.7, cex=0.8)
#mtext("UMAP2", side=2, line=0.7, cex=0.8)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(metadata$Luminal_gr)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T, cex=1)

@@

@@

```{r Suppl. Table 2., include=FALSE}
#saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
top <- top %>% as_tibble(rownames="KO Terms")

pages <- list(list("Across"=top), res) %>% flatten()
write.xlsx(pages, file=paste0("./Suppl.Tbl/","Suppl.Tbl.02",".xlsx"))
```
